Mutual Modelling

### 2015-06-18 authors.tex
Prepa for submission: anonymization + page with authors' information

\keywords{Cognitive Modelling; Grounding; Theory of Mind}

\section{Introduction}

From its very beginning, CSCL research has been following
\citet{roschelle1995construction} suggestion that collaborative learning has
something to do with the process of constructing and maintaining a \emph{shared
understanding} of the task at hand. Building a shared/mutual understanding
refers to the upper class of collaborative learning situations, those in which
students should build upon each other's understanding to refine their own
understanding: what is expected to produce learning is not the mere fact that
two students build the same understanding but the cognitive effort they have to
engage to build this shared understanding~\citep{schwartz1995emergence}. This
effort can be observed by the frequency of rich interactions, \ie interactions
whose occurrence has been related to learning: (self-) explanations in cognitive
science~\citep{chi1989self, webb1991task}, conflict resolution in
socio-cognitive theories~\citep{doise1975social},
mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
construction of a shared understanding has been investigated for several years
in psycholinguistics, under the  notion of
\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
mechanisms for explaining learning outcomes has been questioned in learning
sciences. The monitoring and repair of misunderstanding explains for instance
referential failures in short dialogue episodes but does hardly predict
\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
new belief into one's mental model) over longer
sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
episodes can probably be better understood from a socio-cultural perspective:

\begin{quote}
Collaborative learning is associated with the increased
cognitive-interactional effort involved in the transition from \emph{learning to
understand each other} to \emph{learning to understand the meanings of the semiotic
tools that constitute the mediators of interpersonal
interaction}~\citep{baker1999role}
\end{quote}

Along this line, several scholars suggest that CSCL research should go deeper in
the understanding of how partners engage into shared meaning
making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
making~\citep{suthers2006technology}.

Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
change viewpoint, it is criticized at the same time as being too cognitivist by
some psycholinguists, \ie as overestimating the amount of shared knowledge and
mutual representations actually necessary to conduct a dialogue. The fundamental
issue, as old as philosophy, is the degree of coupling between the different
levels of dialogue, mostly between the lexical/syntactical level and the deeper
semantic levels. \citet{pickering2006alignment} argue that the mutual
understanding starts mostly with a \emph{superficial alignment} at the level of
the linguistic representations, due to priming mechanisms, and that this local
alignment may -- in some cases -- lead to a \emph{global alignment} of the
semantic level (\emph{deep grounding}).  For these authors, the convergence in
dialogue, and even the repair of some misunderstandings, is explained by this
mimetic behavior more than by a monitoring of each other's knowledge:
\emph{``...interlocutors do not need to monitor and develop full common ground
as a regular, constant part of routine conversation, as it would be
unnecessary and far too costly. Establishment of full common ground is, we
argue, a specialized and non-automatic process that is used primarily in
times of difficulty (when radical misalignment becomes apparent).''} (p. 179).
This view is actually not incompatible with Clark's \emph{grounding
criterion}~\citep{clark1989contributing}: the degree of shared understanding that
peers need to reach depends upon the task they perform. For instance, a dialogue
between two surgeons might rely on superficial alignment if they talk about
their friends but has to guarantee accurate common grounds when talking about
which intervention will be conducted in which way on which patient. 
%Some CSCL
%scholars stressed the ``illusion of shared understanding'' \fixme{REF: Jacco?},
%\ie the potential decoupling between linguistic alignment and actual shared
%meanings.

This interesting cognitive science debate occurred mostly outside the field of
learning. In education, the question is to relate these mechanisms to learning
outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
change?} \textbf{Does negotiation of meaning only occurs when partners monitor
and diagnose each other's knowledge?} If the ratio between shallow alignment and
deep grounding depends upon the task, and if deep grounding is a condition for
learning, then the pedagogical challenge is indeed to design tasks that require
deep grounding. Most empirical studies on grounding and alignment are conducted
with tasks, which despite being qualified as ``ecologically valid'' by their
authors, are mere referencing tasks such as asking the way to the train station
or helping the peer to choose a picture among many. In this contribution, we
explore several richer tasks such as arguing on a sensitive issue or building a
concept map.

Deep grounding or shared meaning making requires some cognitive load. For
\citet{clark1986referring}, what is important is not the individual effort made
by the receiver of a communicative act, but the overall \emph{least collaborative
effort}.  The cost of producing a perfect utterance may be higher than the cost
of repairing the problems that may arise through misunderstandings, and in fact,
subjects tend to make less efforts adapting their utterances to a specific
partner when they know that they can later provide feedback on his/her
understanding~\citep{schober1993spatial}. We introduced the notion of
\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
that misunderstanding should not be viewed as something to be avoided (if this
was possible), but as an opportunity to engage into verbalization, explanation,
negotiation, and so forth. This issue is related to the global argument
regarding cognitive load in learning activities, especially in discovery
learning environments: there is no learning without some cognitive load, but
overload may hinder learning~\citep{paas2003cognitive}. In the context of
collaborative learning, we understand the cognitive load induced by mutual
modelling as part of~\citet{schwartz1995emergence} notion of effort towards a
shared understanding. For instance, CSCL researchers expanded the use of
\emph{collaboration scripts}~\citep{kobbe2007specifying}. A script is a pedagogical
method that frames collaborative learning activities in order to foster the
emergence of productive interactions such as argumentation, explanation or
conflict. Conflict-resolution scripts such as the {\sc
ArgueGraph}~\citep{dillenbourg2008mechanics} form pair of students with
opposite opinions, which increases the difficulty of consensus building,
requiring more justifications, more negotiation, more load. Similarly, {\sc
jigsaw} scripts~\citep{aronson1978jigsaw} provide peers with different but
complementary knowledge for augmenting (reasonably) the efforts that group members
have to engage into to reach a shared solution.

%\paragraph{Theory of Mind}
%
%Theory of Mind (originally defined in~\citep{premack1978does}) is the cognitive
%ability that a subject possesses to represent the mental state of another
%agent, possibly including knowledge that contradicts the subject's own model: for
%example, a book can be at the same time \emph{visible} for myself, and \emph{not
%visible} for you.
%
%Children develop this skill, which is essential to understand others'
%perspectives during interactions, around the age of three. It supposes the
%ability to build, store and retrieve separate models of the knowledge of the
%interactors.
%
%One classical application of this cognitive skill is the so-called
%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
%experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
%people, $A$ and $B$, manipulate objects. Then $A$ leaves and $B$ hides away one
%object. When $A$ comes back, we ask the child ``where do you think $A$ will
%look for the object?''. Before acquiring a theory of mind, children are not
%able to separate their own (true) model of the world (where they know that
%the object was hidden) from the model of A, which contains \emph{false
%beliefs} on the world ($A$ still thinks the object is at its original
%position since he did not see $B$ hiding it).

\vspace{2em}
These preliminary remarks frame the main question that this article address:
in order to engage into shared understanding, which is a desirable learning
situation, do partners have to build a representation of each other's knowledge?
if so, what exactly is the nature of this mutual modelling? We adopt hereafter a
constructive approach, first introducing a simple yet convenient model of mutual
modelling, then analyzing five independent studies in terms of the interactions
between learning and mutual modelling.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Mutual Modelling}

We refer to \emph{mutual modelling} as the process of inferring one's partner
mental states. Any claim that students carry out a detailed monitoring of their
peers would be as incorrect as any claim that they do not maintain any
representation at all. If mutual modelling had to be permanently detailed and
accurate, subjects would obviously face a huge cognitive load. Conversely, peers
could not collaborate without some minimal amount of mutual modelling. For
instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
profession) and, of course, it can be incorrect\ldots but it can not remain empty.
Dialogues include many instances of utterances such as ``I thought he would do
that'' (first level of mutual modelling) or even ``He thought I would do that
but I intended something else.'' (second level of mutual modelling).

The content of mutual models ranges from \emph{dispositional} versus
\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
representation of $B$'s long term knowledge, skills or traits. It is thus closely
related to the notion of transactive memory~\citep{wegner1987transactive,
moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
B's knowledge, behavior or intentions specifically activated in the situation in
which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
ones for 2 hours. Examples of fragments that constitute $A$'s model of B
regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:

\begin{itemize}

    \item $Model(A,B, knowledge)$: what does $A$ know about $B$'s knowledge with
        respect to the task at hand or, inversely, about $B$'s knowledge gaps?
        When can $A$ consider $B$'s statements as reliable?

    \item $Model(A,B, skills)$: what does $A$ know about $B$'s skills with respect to
        the task at hand? May $A$ expect $B$ to perform well in a specific subtask?
        The effectiveness of division of labor depends on the quality of this
        mutual model.

    \item $Model(A,B, goals)$: what does $A$ know about $B$'s intentions with respect
        to the project, including $B$' motivation and commitment? Can $A$ trust B
        when $B$ promises to deliver?

    \item $Model(A,B, task)$: what does $A$ know about $B$'s representation of the
        situation and the task: does $A$ knows whether $B$ has the same
        understanding of the problem at stake?

    \item $Model(A,B, plans)$: what does $A$ know about $B$'s strategy. Does A
        understand why $B$ did what he did? Is $A$ able to anticipate what $B$ will do
        next?

    \item $Model(A,B, ``urgent'')$: what does $A$ know about $B$'s understanding of $A$'s
        last utterance: does ``urgent'' means now, soon or ``not too late''?

\end{itemize}

The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
beliefs, emotions, history, status, etc.

Besides, we have different levels of mutual modelling. If $A$ states ``B thinks I am good in
maths'', $A$ builds a \emph{second level} model:
\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
regress of nested models: $A$
saying ``$B$ knows that I don't expect him to solve this statistics problem''
corresponds to \\ \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.

A partner model is likely not a ``box'', \ie not a monolithic representation
but rather a mosaic of information fragments about the partner, with various
granularity and various life cycles. This mosaic is elaborated through a variety
of mechanisms, first for building an initial model of the partner, then for
updating this model. As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other based on cues such as
his/her belonging to broad categories (age, culture, profession, \ldots),
stereotypes (sportsmen, junkie, business women, Swiss,\ldots) as well as physical
appearance. Scholars studied how initial modelling impacts communication. In
their experiments on initial mutual modelling, \citet{slugoski1993attribution}
pretended to their subjects that their (fake) partner had or had not received
the same information. They observed that the subjects adapted their dialogue by
focusing the explanation on the items that he/she was supposed to ignore.
\citet{brennan1991conversation} showed that the subjects used different initial
strategies in forming queries depending on who they were told their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which $A$ and $B$ attended together~\citep{clark2002definite} in the physical
space or in their cultural space (\eg ``09-11''). While co-presence means that
they can refer to shared objects and events, it does not imply that they give
them the same meaning. Namely, a shared screen does not mean a shared
understanding~\citep{dillenbourg2006sharing}.

After initialization, mutual models are updated during the collaborative work
through verbal and non-verbal interactions. A default inference rule is that ``my
partner agrees with me unless he disagrees'', which rejects the critiques that
mutual modelling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing the partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Mutual modelling does not occur in a vacuum but it is highly contextualized.
\citet{clark1991grounding} review how the features of the collaborative
situation, namely the media (co-temporality,\ldots), may facilitate or hamper
mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
short silence between two pilots was perfectly interpreted because it occurred
in a highly constrained communication context.  Some environments are more
productive than others in helping peers to detect their misunderstandings.
\citet{roschelle1995construction} reformulate the design of CSCL
interfaces to provide ways for peers to detect and repair their
misunderstanding. In CSCW, researchers have explored various so-called
``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
that A could not directly perceive because $B$ was working on a different subset
of the virtual space. Different awareness tools are presented and relied upon in
this contribution as well since they allowed us to manipulate experimentally the
mutual modelling activity.

%\subsection{State of the Art}
%
%\paragraph{Building a model of a human during the interaction} Perspective
%Taking is a human ability which allows one to put him/herself in another
%person's point of view. Studied in psychology
%literature~\citep{Flavell1992,Tversky1999}, this ability is crucial when
%interacting with people by allowing one to reason on others' understanding of
%the world in terms of visual perception, spatial descriptions, affordances and
%beliefs, etc.  In the last years these notions have been gradually employed in
%HRI.~\citep{Breazeal2006} presents a learning algorithm that takes into account
%information about a teacher's visual perspective in order to learn a task.
%~\citep{Johnson2005} apply visual perspective taking for action recognition
%between two robots.~\citep{Trafton2005} use both visual and spatial perspective
%taking to find out the referent indicated by a human partner.
%
%The cognitive model that the robot builds for the agent it interacts with is
%still simple and mostly focused on geometric features (who sees what? What are
%our relative positions? etc.). Extending this knowledge with more subtle
%perceptions (emotional state for instance) remains to be done.
%
%\paragraph{Joint action}
%Several theories dealing with collaboration~\citep{Cohen1991,Grosz1996,Clark1996}
%emphasize that collaborative tasks have specific requirements compared to
%individual ones, \eg, since the robot and the person share a common goal, they
%have to agree on the manner to realize it, they must show their commitment to
%the goal during execution, etc. Several robotic systems have already been built
%based on these theories~\citep{Rich1997,Sidner2005,Breazeal2003} and they all
%have shown benefits of this approach. They have also shown how difficult it is
%to manage turn-taking between communication partners and to interleave task
%realization and communication in a generic way. Finally, today only few
%systems~\citep{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
%levels.
%

\section{A model for mutual modelling evaluation}

To report on experiments on mutual modelling, we suggest the notation
\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This is a simple but
useful reduction to reason on mutual modelling. This notation does not mean $A$
has an explicit, monolithic representation of $B$: it must be understood as an
abstraction referring to complex socio-cognitive processes. Besides, we refer to
the \emph{degree of accuracy} of the model as \Model{A}{B}{X}.

We parametrize and assess the mutual modelling effort through 3 variables:

\begin{enumerate}

    \item Tasks vary a lot with respect to how much they require mutual
        understanding.  The \emph{grounding criterion} -- denoted
        \groundingcriterion -- refers to how accurate the models need to be to
        succeed the task $T$. It can be calculated as the probability to succeed
        $T$ despite the fact $X$ is not grounded.
        $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be estimated from the correlation
        between \Model{A}{B}{X} and the task performance.

    \item Before any specific grounding action, there is generally a non-null
        probability that $X$ is mutually understood by $A$ and $B$ (\eg $X$ is
        part of $A$'s and $B$'s cultures, it is manifest to co-present subjects
        or simply there is not much space for misunderstanding or disagreement
        about $X$). We simply could not collaborate without a certain level of
        initial grounds. We note the theoretical accuracy of \emph{initial
        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.

    \item The \emph{cost of grounding} $X$ refers to the physical and cognitive
        effort required to perform a grounding act $\alpha$: a verbal repair
        (\eg rephrasing), a deictic gesture, a physical move to adopt one
        partner's viewpoint, etc. This cost varies according to media
        features~\citep{clark1991grounding}.

\end{enumerate}

Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
content $X$ at time $t_1$ during task T in order to increase \Model{A}{B}{X}
at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
costs (3)~\citep{traum1996miscommunication}:

\begin{multline} \label{eq:probrepair}
    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \simeq
    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) - \mathcal{M}^{\circ}_{t_1}(A,B,X)}{cost (\alpha)}
\end{multline}

This is qualitative summary more that a real equation since several parameters
are hard to quantify (\eg the cost of a communication act depends upon the user
as well) but it clarifies the parameters of the experiments we present hereafter.

\begin{itemize}
    \item \Model{A}{B}{X}: our experiments address different contents that can be
        represented in mutual models:

    \begin{enumerate}

        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
            performed (study 2) or will perform next (study 1),

        \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
            (study 3),

        \item \model{A}{B}{knowledge}: how accurately $A$ estimates $B$'s knowledge
            with respect to the material they learn together (study 4 and 5).

    \end{enumerate}

    \item $\mathcal{M}^{\circ}_{min}(A,B,X,T)$: our studies build upon various
        collaborative tasks: argumentation (study 3), games (study 1 and 2) and
        concept mapping (study 4 and 5). By varying the tasks, we do actually
        vary the grounding criterion but they all require a ``reasonably high''
        grounding criterion.  All tasks are about building a solution or a
        representation together; we did not address everyday conversations.

    \item $\mathcal{M}^{\circ}_{t0}(A,B,X)$: along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4 and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university (and they hence had some general common
        grounds).  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
        using a {\sc jigsaw} script.

    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
        not eliminate all costs: if the tool provide $A$ with information about
        what $B$ does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

\end{itemize}

Measuring \Model{A}{B}{X} is methodologically difficult. We
discriminate two steps, first to capture \model{A}{B}{X} and then to
estimate \Model{A}{B}{X}.

\begin{itemize}
    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$
        what he/she believes about what $B$ knows, feels, intends to do, etc.
        This raises obvious methodological concerns since such a question
        triggers a modelling process beyond what would naturally occur. To avoid
        this bias, one can estimate mutual modelling after task completion.
        Then, the obvious drawback are memory losses and by post-hoc
        reconstruction.  The first method was used in study 2 and the second one
        in the other studies. Another option would be to rely on external
        behavioural metrics like eye-tracking: we hypothesise for instance that
        the fixation time reflects the efforts engaged by the human to
        understand, hence, model, the others. Such an approach has however not
        been investigated in the presented studies.

    \item {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
        need to access the reference model \refmodel{B}{X} to estimate its
        \emph{accuracy}. Since we can only indirectly access it via what $B$
        reports (\ie \model{B}{B}{X}), accuracy can be estimated in 2 ways:

        \begin{itemize}

            \item Subjective accuracy: In study 3, we compute
                \Model{A}{B}{X} by measuring if $A$ describes
                $B$'s emotions in the same way $B$ reports its emotions 
                ($\M{A}{B}{X} = \M{B}{B}{X}$).

            \item Objective accuracy: In studies 4 and 5, we compute
                \Model{A}{B}{X} by comparing \Model{A}{B}{K} to $B$'s
                actual knowledge as it has measured by a test.

        \end{itemize}

\end{itemize}

\subsubsection*{Hypotheses}

The experiments we report here address mutual modelling across different tasks,
some with dyads, other with triads. They were conducted along 6 years in two
different institutions by different researchers. They used different
independent, intermediate and dependent variables. Nonetheless, we were able to
address a few questions across these studies. We start with 3 simple hypotheses
about $\mathcal{M}^{\circ}(A,B)$:

\begin{itemize}
    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon $A$'s ability or effort
        to model $B$,

    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability or
        effort to help $A$ to model him/herself,

    \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
        interactions among $A$ and $B$.

\end{itemize}

$\mathcal{H}_{2}$ relates to second level modelling since $B$ needs to monitor
$A$ to see if $A$ understood him/her ($\mathcal{M}(B,A,\mathcal{M}(A,B))$). We
will see that $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ are actually difficult to
differentiate.

\paragraph{The symmetry question}

One interesting question relates to the symmetry of mutual modelling: what is
the relationship between \Model{A}{B}{X} and \Model{B}{A}{X}
(Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is
mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
might support $\mathcal{H}_{2}$ or $\mathcal{H}_{3}$ since there is a low
probability that randomly formed pairs integrate peers with the same level of
mutual modelling skills.

\begin{figure}[htb]
\centering

\begin{tikzpicture}[>=latex, scale=0.5]

\draw(0,0) node[anchor=north] (A) {\scriptsize A};
\draw(10,0) node[anchor=north] (B) {\scriptsize B};
\draw(5,2) node[anchor=north] {\scriptsize \Model{A}{ B}{ X}};
\draw(5,-2) node[anchor=north] {\scriptsize \Model{B}{ A}{ X}};
\draw[dashed, <->] (5,1) -- (5,-1.9) node[midway, sloped, above] {$\Delta_1$};
\draw[->] (A) to[bend left] (B);
\draw[->] (B) to[bend left] (A);

\end{tikzpicture}

\caption{\small Mutual modelling in a dyadic interaction, $\Delta_1 =
    \Delta(\mathcal{M}^{\circ} (A,B,X),
\mathcal{M}^{\circ} (B,A,X))$}

\label{mm_symmetry}
\end{figure}

\paragraph{The triangle questions}

With triads, we may compute the accuracy of 6 models:
\Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
\Model{C}{B}{X} and \Model{B}{C}{X}.

This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
and $B$ have the same accuracy when modelling $C$ ($\Delta_2 =
\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? If it is the
same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gains over $\mathcal{H}_{1}$

Conversely, does $C$ model more accurately $A$ than $B$ ($\Delta_3=
\Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
answer would support $\mathcal{H}_{3}$. A negative answer would support
$\mathcal{H}_{2}$ since $\Delta_3$ could mainly be explained by the way $A$ and
$B$ help $C$ to model them.

\begin{figure}[htb]
\centering
\subcaptionbox{}{ 
    \begin{tikzpicture}[>=latex, scale=0.5]

    \draw(0,2) node (A) {\scriptsize A};
    \draw(0,-2) node (B) {\scriptsize B};
    \draw(10,0) node (C) {\scriptsize C};

    \draw(5,2) node {\scriptsize \Model{A}{ C}{ X}};
    \draw(5,-2) node {\scriptsize \Model{B}{ C}{ X}};
    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above] {$\Delta_2$};
    \draw[->] (A) to (C);
    \draw[->] (B) to (C);

    \end{tikzpicture}
}
\subcaptionbox{}{ 
    \begin{tikzpicture}[>=latex, scale=0.5]

    \draw(0,2) node (A) {\scriptsize A};
    \draw(0,-2) node (B) {\scriptsize B};
    \draw(10,0) node (C) {\scriptsize C};

    \draw(5,2) node {\scriptsize \Model{C}{ A}{ X}};
    \draw(5,-2) node {\scriptsize \Model{C}{ B}{ X}};
    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above]
    {$\Delta_3$};
    \draw[<-] (A) to (C);
    \draw[<-] (B) to (C);

    \end{tikzpicture}
}
\caption{\small Mutual modelling in a triadic interaction}

\label{mm_triangles}
\end{figure}

In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
whether the accuracy of mutual modelling depends more upon the modeller's effort
or the modellee's behaviour.

\paragraph{The rectangle questions}

We can go further by comparing self- versus other modelling ($\Delta_4$ in
Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
question if modelling skills depend upon what aspects are being modeled ($X$ or
$Y$), which would explain vertical differences ($\Delta_5$ in
Fig.~\ref{mm_rectangle}). When possible, we will also discuss these two
questions in the studies presented hereafter.

\begin{figure}[htb]
\centering

\begin{tikzpicture}[>=latex, scale=0.5]

    \draw(0,0) node (a) {\scriptsize \Model{A}{ B}{ X}};
    \draw(10,0) node (b) {\scriptsize \Model{A}{ A}{ X}};
    \draw(10,-4) node (c) {\scriptsize \Model{A}{ A}{ Y}};
    \draw(0,-4) node (d) {\scriptsize \Model{A}{ B}{ Y}};
    \draw[<->] (a) -- (b) node[midway, below] {$\Delta_4$};
    \draw[<->] (b) -- (c);
    \draw[<->] (c) -- (d);
    \draw[<->] (d) -- (a) node[midway, sloped, above] {$\Delta_5$};

\end{tikzpicture}

\caption{\small Meta-cognitive skills and domain-dependent modelling}

\label{mm_rectangle}
\end{figure}

This simple notation does not pretend to provide a mathematical account of
mutual modelling but to be more systematic in describing the following
experiments.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Studies}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 1}: Effect of an awareness tool on \gModel{A}{B} in a virtual
game}

We studied the impact of an awareness tool on group performance and mutual
modelling~\citep{nova2007collaboration}. The availability
of an awareness tool was our independent variable. In previous studies, we
replayed a video of the game to subjects who surprised us by their ability to
remember former states of their mutual model: ``I did that because I thought that
you would do that''. Hence, this experiment focused the representation of each
other's action plans. During the game, we asked them to anticipate the next
action of their partner as well as to announce their own actions.

\subsubsection*{Experimental setting}

\begin{figure*}
        \centering
        \includegraphics[width=\textwidth]{image4.png}
        \caption{Screenshot of the {\sc SpaceMiners} game}
        \label{study1:spaceminer}
\end{figure*}

{\sc SpaceMiners} is a 3D game that involves two players collecting minerals located
in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch drones through
the space after choosing their initial direction and speed. Once launched, the
trajectory of drones is only influenced by the gravity of planets and by
trajectory modification tools.  During the experiment, the teams were confronted
with three increasingly complex situations. The experiment was 2 hours long: a
30 minutes tutorial and 3 levels of 30 minutes. Players uses using a regular
joystick and communicated with each other through an audio channel.

The independent variable was the availability of an awareness tool that shows to
player $A$  the location and gaze direction of player $B$. In the awareness
condition, players could switch to the \emph{scout mode} where they could view what
their partner was looking at. We hypothesize that this would enable subjects to
more accurately infer their teammate's intentions. Each player sat in front of a
distinct computer located in different rooms.

\subsubsection*{Subjects}

Thirty-six persons participated in this study, all native French speakers. We
constituted 18 dyads who did not know each other beforehand. The pairs were
randomly assigned to either the control condition (without the awareness tool)
or the awareness condition (with the awareness tool).

\subsubsection*{Variables}

Task performance was measured by the score reached by the two subjects at the
end of the game (three levels). The effort of mutual modelling was measured as
the ratio of time that players would spend in the scout mode (divided by total
time), which is the time during which players are not performing their own
actions but monitoring their partner's actions.

In order to evaluate \gModel{A}{B} during the task, we used two questionnaires
(Fig.~\ref{study1:questionnaires}) that were displayed during each of the three
games, as a transparent layer appearing over the game display. The first
questionnaire concerned the player's intended actions. The second questionnaire
asked each player about what he thought her or his partner was intending to do.
Some answers were identical in both questionnaires (like ``adjusting a shot'')
while others were reversed (``to guide my partner'' versus ``to guide me'').
This method provides us with a subjective measure of accuracy
($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$) rather than an objective measure (\ie the
model \gmodel{A}{B} is compared to $B$'s next action) because some of the
activities proposed by the questionnaire were not observable by the environment
(\eg establishing a strategy). We calculated \Model{A}{B}{activity})
as the number of common answers between questionnaires \gmodel{A}{A} and
\gmodel{A}{B} in each game and computed the average value a across the 3 levels.

\begin{figure}[ht!]
        \centering
        \includegraphics[width=0.7\columnwidth]{image5.png}
        \caption{\gmodel{A}{A} and \gmodel{A}{B} questionnaires in SpaceMiners
        (translated from French).}

        \label{study1:questionnaires}
\end{figure}

\subsubsection*{Results}

\paragraph{Grounding criterion} The grounding criterion was high: the
correlation between \gModel{A}{B} and task performance was 0.42, $p = 0.05$.
Pairs with an accurate mutual model reached higher scores. A regression analysis
confirmed the positive and significant relation between group performance and
mutual modelling accuracy ($\beta=54, p = 0.02$).

\paragraph{Study-specific questions} The awareness tool permitted higher group
performance, but it did not improve the accuracy of the mutual model. Since
teams were free to use the awareness tool or not (the \emph{scout} mode), we
performed a post-hoc split of players depending on how much time they used it.
The split point was the mean of time spent in the scout mode and it led to the
constitution of two groups made up of 12 individuals ``short time in scout
mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
variance conducted on these contrasted groups revealed that pairs in the
awareness condition who spent more time in the scout mode reached higher levels
of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
does not support a causal direction. An alternative explanation could be that
good modellers are more social and hence appreciate the awareness tool.

\paragraph{Symmetry question} We computed intra-class correlation as described
by~\citet{kenny1998data} from the answers to the cross-questionnaires.
Considering all pairs in both conditions, we found a positive and significant
correlation ($r = 0.38, p < 0.05$) between \gModel{A}{B} and \gModel{B}{A}.
Interestingly, this was higher in the control group ($r = 0.44$) than in the
experimental group ($r = 0.24$). Actually, $\Delta(\gMdeg{A}{B},\gMdeg{B}{A})$,
\ie the absolute differences between the models accuracy, was not significantly
different with or without the awareness tool ($F [1,13]= 0.144, p > 0.5$). This
result could be explained by the fact that the players without awareness tools
communicated more.

\subsubsection*{The triangle and rectangle questions are not addressed in this study.}

\subsubsection*{Discussion}

How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
If it was null, modelling would be interpreted as an individual process that
varies according to personality traits or cognitive skills: some persons tend to
model more accurately their peer while others don't care or can simply not
($\mathcal{H}_{1}$). If the correlation was close to 1, we could conclude that
the modelling accuracy is an emergent property of teams since there would be low
probability that, by chance, low accurate modellers are paired with accurate
modellers and conversely ($\mathcal{H}_{2}$). Our correlation is between these
two extremes but the very notion of significant intra-class correlation tend to
indicate that one may not consider peers as independent subjects, which is an
argument for the \emph{interaction} hypothesis $\mathcal{H}_{3}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{{\bf Study 2}: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
game}

This study concerns a collaborative game that occurred in the physical
space~\citep{nova2006underwhelming}. We studied whether players build an accurate
model of the path followed by their partners, assuming this path reflected their
problem solving strategy. We used an objective measure of \gModel{A}{B}: the
distance between where $A$ believes $B$ has been walking and where $B$ actually went.
The main hypothesis concerned the effect of awareness tools on group performance
and on \gModel{A}{B}.

\subsubsection*{Experimental setting}

>>>>

\subsection*{Corresponding author}
Séverin Lemaignan \texttt{severin.lemaignan@epfl.ch}
\end{document}
